---
title: "intro to viz"
author: "Charles Lang"
date: "October 3, 2016"
output: html_document
---
#Input
```{r}
D1 <- read.csv("School_Demographics_and_Accountability_Snapshot_2006-2012.csv", header = TRUE, sep = ",")

D2 <- subset(D1, schoolyear == 20112012)
```

#Histograms
```{r}
hist()

hist(D2$frl_percent, breaks = 100)

hist(D2$frl_percent, breaks = c(0,10,20,80,100))

hist(D2$frl_percent, breaks = 100, ylim = c(0,30))
```

#Plots
```{r}
plot(D2$ell_num, D2$ctt_num)

plot(D2$ell_num ~ D2$ctt_num)

#Barplot
x <- c(1,3,2,7,6,4,4)
y <- c(2,4,2,3,2,4,3)

table1 <- table(x,y)

barplot(table1)

#Lineplot
D3 <- aggregate(D1, by = list(D1$schoolyear), FUN = mean)

plot(D3$schoolyear, D3$total_enrollment, type = "l", lty = "dashed")

#Boxplot
D5 <- subset(D1, DBN == "31R075")
D5 <- droplevels(D5)
boxplot(D5$total_enrollment ~ D5$Name)
```
#Pairs
```{r}
D4 <- D2[,c(5,6, 21:24)]

pairs(D4)
```
# Exercise

1. Create a simulated data set containing 100 students, each with a score from 1-100 representing performance in an educational game. The scores should tend to cluster around 75. Also, each student should be given a classification that reflects one of four interest groups: sport, music, nature, literature.

```{r}
#rnorm(100, 75, 15) creates a random sample with a mean of 75 and standard deviation of 20
#pmax sets a maximum value, pmin sets a minimum value
#round rounds numbers to whole number values
#sample draws a random samples from the groups vector according to a uniform distribution

scores <- round(pmax(1, pmin(100,rnorm(100, 75, 20))))
studentid <- seq(1,100,1)
groups <- c("sport", "music", "nature", "literature")
interest <- sample(groups, 100, replace = TRUE)
D1 <- data.frame(studentid, scores, interest)

```

2. Using base R commands, draw a histogram of the scores. Change the breaks in your histogram until you think they best represent your data.

```{r}
hist(D1$scores, breaks = 10)
```


3. Create a new variable that groups the scores according to the breaks in your histogram.

```{r}
#cut() divides the range of scores into intervals and codes the values in scores according to which interval they fall. We use a vector called `letters` as the labels, `letters` is a vector made up of the letters of the alphabet.
D1$binned = cut(D1$scores, breaks=10, labels = letters[1:10])
```

4. Now using the colorbrewer package (RColorBrewer; http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) design a pallette and assign it to the groups in your data on the histogram.

```{r}
library(RColorBrewer)
#Let's look at the available palettes in RColorBrewer
display.brewer.all()
#The top section of palettes are sequential, the middle section are qualitative, and the lower section are diverging.
#Make RColorBrewer palette available to R and assign to your bins
D1$colors <- brewer.pal(10, "Set3")
#Use named palette in histogram
hist(D1$score, col = D1$colors)
```


5. Create a boxplot that visualizes the scores for each interest group and color each interest group a different color.

```{r}
#Make a vector of the colors from RColorBrewer
interest.col <- brewer.pal(4, "Dark2")

boxplot(scores ~ interest, D1, col = interest.col)

plot(D1$interest, D1$scores)
```


6. Now simulate a new variable that describes the number of logins that students made to the educational game. They should vary from 1-25.

```{r}
D1$login <- sample(1:25, 100, replace = TRUE)
```

7. Plot the relationships between logins and scores. Give the plot a title and color the dots according to interest group.

```{r}
plot(D1$login, D1$scores, col = D1$col1, main = "Student Logins vs. Scores")

D1$col1 <- ifelse(D1$interest == "music", "red", "green")

```


8. R contains several inbuilt data sets, one of these in called AirPassengers. Plot a line graph of the the airline passengers over time using this data set.

```{r}
AP <- data.frame(AirPassengers)
plot(AirPassengers)
```


9. Using another inbuilt data set, iris, plot the relationships between all of the variables in the data set. Which of these relationships is it appropraiet to run a correlation on? 

```{r}
pairs(iris)

#Correlations characterize linear relationships.
```

